Auditory Scene Similarity Retrieval and Classification with Relevance-based Quantization of Scattering Features

### 2016-12-14 paper/structureScene.tex
TODO on asymmetry of masking in Gammatones

\subsection{Gammatone wavelets}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{bw/gammatones}
\caption{
\label{fig:gammatones}
Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$. 
Oscillations represent the real and imaginary parts. The orange envelope represents the complex modulus.}
\end{center}
\end{figure}
Wavelets
$\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as fourth-order Gammatone
wavelets with one vanishing moment \cite{Venkitaraman2014}, and are shown in Figure \ref{fig:gammatones}.
In the context of auditory scene analysis, the asymmetric envelope of Gammatone wavelets is more biologically plausible than the symmetric, Gaussian-like envelope of the more widely used Morlet wavelets.
Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking  \cite{Fastl2007}.
Moreover, it should be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
As such, they are similar to filters discovered automatically by unsupervised encoding of natural sounds \cite{Smith2006}.
This suggests that, despite being hand-crafted and not learned, Gammatone wavelets provide a sparser time-frequency representation of acoustic scenes compared to other variants.

>>>>

\subsection{Gammatone wavelets}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{bw/gammatones}
\caption{
\label{fig:gammatones}
Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$. 
Oscillations represent the real and imaginary parts. The orange envelope represents the complex modulus.}
\end{center}
\end{figure}
Wavelets
$\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as fourth-order Gammatone
wavelets with one vanishing moment \cite{Venkitaraman2014}, and are shown in Figure \ref{fig:gammatones}.
In the context of auditory scene analysis, the asymmetric envelope of Gammatone wavelets is more biologically plausible than the symmetric, Gaussian-like envelope of the more widely used Morlet wavelets.
Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking  \cite{Fastl2007}.\vl{TODO}
Moreover, it should be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
As such, they are similar to filters discovered automatically by unsupervised encoding of natural sounds \cite{Smith2006}.
This suggests that, despite being hand-crafted and not learned, Gammatone wavelets provide a sparser time-frequency representation of acoustic scenes compared to other variants.
